\Title

### 2021-06-28 manuscript/content.tex
Clarify the part about the A_k Jacobian matrices
Break apart the item that describes them to highlight the definition.

\begin{enumerate}
  \item Define a set of $M$ equivalent sources distributed throughout the
    survey area (see section \ref{sec:source_distribution} for details).
  \item Define a set of $K$ overlapping windows of equal size that cover the
    survey area.
  \item Create $K$ separate sets of equivalent sources, one for each window.
    Each set will be formed by the portion of the original $M$ sources that
    fall inside the respective window.
    Since the windows overlap, the total number of sources from all sets will
    be greater than $M$.
  \item Define vector $\mathbf{c}_k$ as the coefficients of the equivalent
    sources of the $k$-th window.
  \item Define matrix $\mathbf{A}_k$ as the Jacobian matrix between the sources
    in the $k$-th window and all $N$ data points of the survey.
  \item Model the predicted data as a superposition of the effects of the $K$
    separate sets of equivalent sources (i.e., Eq.~\ref{eq:gb-linear-model}).
\end{enumerate}

The gradient boosting algorithm works by fitting each component of the
additive model, one at a time, to the residuals of the previous component.
\citet{friedman2001} demonstrates that this corresponds to a steepest-descent
optimization in the so-called ``function space''.
The adaptation of the gradient boosting method to find the damped least-squares
solutions for the $K$ parameters vectors $\mathbf{c}_k$ in
Eq.~\ref{eq:gb-linear-model} is presented in
Algorithm~\ref{alg:gradient_boosting}.

\begin{algorithm}[!h]
  \DontPrintSemicolon
  \setstretch{1.5}
  define the residual vector: $\mathbf{r}_{0} = \mathbf{d}^o$ \;
  \For{ $k = 1$ \KwTo $K$ }{

    calculate $N \times M_k$ Jacobian matrix $\mathbf{A}_k$
    \;

@@

\begin{equation}
    d(\mathbf{p}) =
    \sum\limits_{k=1}^K \sum\limits_{j=1}^{M_k}
    \frac{{c_k}_j}{\left\lVert \mathbf{p} - \mathbf{q}_j \right\rVert}
    \ ,
\end{equation}

\noindent where ${c_k}_j$ is the $j$-th element of the $\mathbf{c}_k$ vector.

To improve the convergence of the algorithm, \citet{friedman2002} suggests
introducing randomness into the fitting process. We achieve this by randomizing
the order in which the $K$ windows are used in the gradient boosting algorithm.

@@

We can improve the efficiency of the algorithm further by using only the $N_k$ data
points that fall within the $k$-th window for fitting the sources
(steps \ref{alg:scale} and \ref{alg:fit} of
algorithm~\ref{alg:gradient_boosting}),
while using all available data when calculating the step size and residuals
(steps \ref{alg:predicted} and \ref{alg:residual}
of algorithm~\ref{alg:gradient_boosting}).
By doing so, we can reduce the size of the Jacobian matrix further to $N_k
\times M_k$ elements.
The forward modeling operation performed in step \ref{alg:predicted} can be
done by a summation (Eq.~\ref{eq:forward-sum})
instead of a matrix-vector product, which allows us to avoid computing and
storing the larger $N \times M_k$ Jacobian at any point.
Algorithm~\ref{alg:gradient_boosting_window} incorporates these changes and is
the final \textit{gradient-boosted equivalent source algorithm}.
Figure~\ref{fig:gradient-boosting-schematics} shows a sketch of the
gradient-boosted equivalent source algorithm with a set of observation points
that simulate a ground survey and locating one source below each data point.

\begin{algorithm}[!h]
  \DontPrintSemicolon
  \setstretch{1.5}
  define the residual vector: $\mathbf{r}_{0} = \mathbf{d}^o$ \;
  \For{ $k = 1$ \KwTo $K$ }{

    select weights $\tilde{\mathbf{W}}_k$ and residuals
    $\tilde{\mathbf{r}}_{k - 1}$ for data points inside the $k$-th window only
    \;

    calculate $N_k \times M_k$ Jacobian matrix $\tilde{\mathbf{A}}_k$
    with data points and sources inside the $k$-th window only
    \;

@@

    calculate $ \mathbf{d}_k $, where
    $
    {d_k}_i
    =
    \sum\limits_{j=1}^{M_k} \dfrac{{c_k}_j}{\left\lVert \mathbf{p}_i -
        {\mathbf{q}_k}_j
    \right\rVert}
    \ \forall i=1\ \text{to}\ N
    $
    \;

    $\mathbf{r}_k = \mathbf{r}_{k - 1} - \mathbf{d}_k$
    \;
  }
  \BlankLine
  \setstretch{1}
  \caption{Gradient-boosted equivalent source algorithm.}
  \label{alg:gradient_boosting_window}
\end{algorithm}

@@

It is worth noting that two sets of equivalent sources obtained through two
adjacent overlapping windows have some portion of the sources on the same
locations, specifically the ones that fall on the intersection between the two
windows.
We can interpret this as the gradient-boosting algorithm fitting the source
coefficients multiple times: one time for every window that covers each source.
This fact can be exploited in order to save some computational memory.
Instead of storing all of the multiple $\mathbf{c}_k$ vectors
(Eq.~\ref{eq:gb-linear-model}), we can initialize a single $\mathbf{c}$ vector
with zeros, where each element represents the coefficient of each one of the
original $M$ sources.
After each iteration of the gradient-boosting algorithm, we add the estimated
coefficients $\hat{\mathbf{c}}_k$ to the corresponding elements of vector
$\mathbf{c}$.
Because the relation between the source coefficients and the effect they
generate is linear, we can safely compute the resulting field through
Equation~\ref{eq:eql-forward}.
This way, the memory needed to store the entire set of estimated coefficients
is limited to a single vector of $M$ elements.

@@

\begin{enumerate}
  \item
    \emph{Sources below data points}: one equivalent source is placed at the
    horizontal location of each data point (Fig.~\ref{fig:source_layouts}b).
    Therefore, the number of sources is be equal to the number of observations
    ($N=M$).
  \item
    \emph{Regular grid}: a homogeneous distribution of point sources below the
    survey region (Fig.~\ref{fig:source_layouts}c). A padding region is often
    added to help reduce edge effects. In practice, it often leads to
    underdetermined problems since a large number of sources is required
    ($N<M$).
\end{enumerate}

@@

\begin{enumerate}
    \item Dividing the survey region into rectangular blocks of equal size.
    \item \label{item:median-position} Computing the median horizontal position
        of the observation points that fall inside each block. Blocks without
        any observation point are omitted.
    \item Assign one point source to the median horizontal position calculated
        in step \ref{item:median-position}.
\end{enumerate}

The number of sources created by this new layout will be less than the number
of observations if the block size is chosen appropriately (i.e., making sure
that blocks are large enough to contain more than a single data point).
The overdetermined problem that arises from this layout has a lower
computational load and is less prone to overfitting the data since the model
complexity is lower.
Moreover, the block averaging process can balance the spacing between sources
along a flight line and between adjacent lines, helping to reduce aliasing
effects in generated grids.
We will demonstrate through tests on synthetic data that the block-averaged sources
layout is able to interpolate with comparable accuracy to other layouts while
using a fraction of the equivalent sources.

@@

\begin{figure*}
    \includegraphics[width=\linewidth]{figs/depth_types.pdf}
    \caption{
        Examples of different strategies for assigning depths to equivalent
        sources.
        Here we assign one source for each
        observation point, located at the same horizontal coordinates as the
        data points.
        Source depths are
        (a)~a \emph{constant depth} at a chosen vertical coordinate,
        (b)~a \emph{relative depth} determined by uniformly downward shifting
        the vertical coordinate of data points,
        and
        (c)~a \emph{variable depth} determined by shifting the vertical
        coordinates of the observation points by an amount proportional to the
        average distance to neighbouring sources.
        The distance between data points and their respective sources (a)
        depends on observation height, (b) is constant, and (c) is proportional
        to the horizontal distribution of sources.
        Notice how the closely spaced sources in the middle of the profile (c)
        are shallower than their counterparts in (b).
    }
    \label{fig:depth_types}
\end{figure*}

@@

\noindent
in which $z$ is the depth of an equivalent source,
$\Delta z$ is a relative depth shift that is the same for all sources,
$\alpha$ is an adimensional depth factor,
$h$ is the median horizontal distance to the $k$ nearest neighbouring sources,
and
$z_{obs}$ is a vertical observation coordinate that will depend on the
horizontal layout strategy.
For \emph{sources below data}, it is the vertical coordinate of the data point
corresponding to the given source.
For \emph{regular grid}, it can be interpolated from the vertical coordinates
of all data points.
Finally, for \emph{block-averaged sources} it will be the average vertical
coordinate of data within the corresponding block.

In the following sections, we will test the effectiveness and trade-offs of
each of these strategies on synthetic data.

@@

\subsection{
    Interpolating synthetic survey with gradient boosting
}

>>>>

\begin{enumerate}
  \item Define a set of $M$ equivalent sources distributed throughout the
    survey area (see section \ref{sec:source_distribution} for details).
  \item Define a set of $K$ overlapping windows of equal size that cover the
    survey area.
  \item Create $K$ separate sets of equivalent sources, one for each window.
    Each set will be formed by the portion of the original $M$ sources that
    fall inside the respective window.
    Since the windows overlap, the total number of sources from all sets will
    be greater than $M$.
  \item Define vector $\mathbf{c}_k$ as the $M_k$ coefficients of the
    equivalent sources of the $k$-th window.
  \item Define matrix $\mathbf{A}_k$ as the $N \times M_k$ Jacobian matrix
    between the sources in the $k$-th window and all $N$ data points of the
    survey.
  \item Model the predicted data as a superposition of the effects of the $K$
    separate sets of equivalent sources (i.e., Eq.~\ref{eq:gb-linear-model}).
\end{enumerate}

The gradient boosting algorithm works by fitting each component of the
additive model, one at a time, to the residuals of the previous component.
\citet{friedman2001} demonstrates that this corresponds to a steepest-descent
optimization in the so-called ``function space''.
The adaptation of the gradient boosting method to find the damped least-squares
solutions for the $K$ parameter vectors $\mathbf{c}_k$ in
Eq.~\ref{eq:gb-linear-model} is presented in
Algorithm~\ref{alg:gradient_boosting}.

\begin{algorithm}[!h]
  \DontPrintSemicolon
  \setstretch{1.5}
  Define the residual vector $\mathbf{r}_{0} = \mathbf{d}^o$ \;
  \For{ $k = 1$ \KwTo $K$ }{

    Calculate the $N \times M_k$ Jacobian matrix $\mathbf{A}_k$
    \;

@@

\begin{equation}
    d(\mathbf{p}) =
    \sum\limits_{k=1}^K \sum\limits_{j=1}^{M_k}
    \frac{{c_k}_j}{\left\lVert \mathbf{p} - {\mathbf{q}_k}_j \right\rVert}
    \ ,
    \label{eq:eql-forward-gb}
\end{equation}

\noindent
in which ${c_k}_j$ is the $j$-th element of the $\mathbf{c}_k$ vector and the
${\mathbf{q}_k}_j$ is the position vector of the $j$-th source of the $k$-th
window.

To improve the convergence of the algorithm, \citet{friedman2002} suggests
introducing randomness into the fitting process. We achieve this by randomizing
the order in which the $K$ windows are used in the gradient boosting algorithm.
Section~\ref{sec:gb_interpolation} explores the effect of randomization in the
convergence rate of the algorithm and the accuracy of the interpolation.

@@

We can improve the efficiency of the algorithm further by:

\begin{enumerate}
  \item Using only the $N_k$ data points that fall within the $k$-th window for
    fitting the sources (steps \ref{alg:scale} and \ref{alg:fit} of
    algorithm~\ref{alg:gradient_boosting}).
    By doing so, we can replace the $N \times M_k$ Jacobian matrix $\mathbf{A}_k$
    with the smaller $N_k \times M_k$ matrix $\tilde{\mathbf{A}}_k$.
    We still use all $N$ data points when calculating the predicted data and
    residuals (steps \ref{alg:predicted} and \ref{alg:residual} of
    algorithm~\ref{alg:gradient_boosting}).
  \item The forward modeling operation performed in step \ref{alg:predicted}
    can be done by a summation (Eq.~\ref{eq:forward-sum}) instead of a
    matrix-vector product, which allows us to avoid computing and storing the
    larger $N \times M_k$ matrix $\mathbf{A}_k$ at any point.
\end{enumerate}

Algorithm~\ref{alg:gradient_boosting_window} is the final
\textit{gradient-boosted equivalent sources algorithm} which incorporates these
changes.
Figure~\ref{fig:gradient-boosting-schematics} shows a sketch of the algorithm
steps applied a set of observation points that simulate a ground survey and
locating one source below each data point.

\begin{algorithm}[!h]
  \DontPrintSemicolon
  \setstretch{1.5}
  Define the residual vector $\mathbf{r}_{0} = \mathbf{d}^o$ \;
  \For{ $k = 1$ \KwTo $K$ }{

    Select weights $\tilde{\mathbf{W}}_k$ and residuals
    $\tilde{\mathbf{r}}_{k - 1}$ for data points inside the $k$-th window
    \;

    Calculate Jacobian matrix $\tilde{\mathbf{A}}_k$ with data points and
    sources inside the $k$-th window
    \;

@@

    Calculate $ \mathbf{d}_k $, where
    $
    {d_k}_i
    =
    \sum\limits_{j=1}^{M_k} \dfrac{{c_k}_j}{\left\lVert \mathbf{p}_i -
        {\mathbf{q}_k}_j
    \right\rVert}
    \quad \forall\ i=1\ \text{to}\ N
    $
    \;

    $\mathbf{r}_k = \mathbf{r}_{k - 1} - \mathbf{d}_k$
    \;
  }
  \BlankLine
  \setstretch{1}
  \caption{Gradient-boosted equivalent sources algorithm.}
  \label{alg:gradient_boosting_window}
\end{algorithm}

@@

It is worth noting that two sets of equivalent sources obtained through two
adjacent overlapping windows have some portion of the sources on the same
locations, specifically the ones that fall on the intersection between the two
windows.
We can interpret this as the gradient-boosting algorithm fitting the source
coefficients multiple times: one time for every window that covers each source.
This fact can be exploited in order to save computer memory.
Instead of storing all of the $\mathbf{c}_k$ vectors
(Eq.~\ref{eq:gb-linear-model}), we can initialize a single $\mathbf{c}$ vector
with zeros, where each element represents the coefficient of each one of the
original $M$ sources.
After each iteration of the gradient-boosting algorithm, we add the estimated
coefficients $\hat{\mathbf{c}}_k$ to the corresponding elements of vector
$\mathbf{c}$.
Because the forward modelling function is linear, we can safely compute the
resulting field through Eq.~\ref{eq:eql-forward} instead of
Eq.~\ref{eq:eql-forward-gb}.
This way, the memory needed to store the entire set of estimated coefficients
is limited to a single vector of $M$ elements.

@@

\begin{enumerate}
  \item
    \emph{Sources below data points}: one equivalent source is placed at the
    horizontal location of each data point (Fig.~\ref{fig:source_layouts}b).
    Therefore, the number of sources is equal to the number of observations
    ($M=N$).
  \item
    \emph{Regular grid}: a homogeneous distribution of point sources below the
    survey region (Fig.~\ref{fig:source_layouts}c). A padding region is often
    added to help reduce edge effects. In practice, it often leads to
    underdetermined problems since a large number of sources is required
    ($M>N$).
\end{enumerate}

@@

\begin{enumerate}
    \item Dividing the survey region into rectangular blocks of equal size.
    \item \label{item:median-position} Computing the median horizontal position
        of the observation points that fall inside each block. Blocks without
        any observation point are omitted.
    \item Assign one point source to each of the median horizontal positions
      calculated in step \ref{item:median-position}.
\end{enumerate}

The number of sources created by this new layout will be less than the number
of observations if the block size is chosen appropriately (i.e., making sure
that blocks are large enough to contain more than a single data point).
The overdetermined problem that arises from this layout has a lower
computational load and is less prone to overfitting the data since the model
complexity is lower.
Moreover, the block averaging process can balance the spacing between sources
along a flight line and between adjacent lines, helping to reduce aliasing
effects in the generated grids.
In Section~\ref{sec:synthetic_distributions}, we demonstrate through tests on
synthetic data that the block-averaged sources layout is able to interpolate
with comparable accuracy to other layouts while using a fraction of the
equivalent sources.

@@

\begin{figure*}
    \includegraphics[width=\linewidth]{figs/depth_types.pdf}
    \caption{
        Examples of different strategies for assigning depths to equivalent
        sources.
        Here we assign one source for each
        observation point, located at the same horizontal coordinates as the
        data points.
        Source depths are
        (a)~a \emph{constant depth} at a chosen vertical coordinate,
        (b)~a \emph{relative depth} determined by uniformly shifting downward
        the vertical coordinate of data points,
        and
        (c)~a \emph{variable depth} determined by shifting the vertical
        coordinates of the observation points by an amount proportional to the
        average distance to neighbouring sources.
        The distance between data points and their respective sources (a)
        depends on observation height, (b) is constant, and (c) is proportional
        to the horizontal distribution of sources.
        Notice how the closely spaced sources in the middle of the profile (c)
        are shallower than their counterparts in (b).
    }
    \label{fig:depth_types}
\end{figure*}

@@

\noindent
in which $z$ is the vertical coordinate (positive downwards) of an equivalent
source,
$\Delta z$ is a relative depth shift that is the same for all sources,
$\alpha$ is an adimensional depth factor,
$h$ is the median horizontal distance to the $k$ nearest neighbouring sources,
and
$z_{obs}$ is a vertical observation coordinate that will depend on the
horizontal layout strategy.
For \emph{sources below data}, it is the vertical coordinate of the data point
corresponding to the given source.
For \emph{regular grid}, it can be interpolated from the vertical coordinates
of all data points.
Finally, for \emph{block-averaged sources} it will be the median vertical
coordinate of the data within the corresponding block.

In Section~\ref{sec:synthetic_distributions}, we test the effectiveness each of
these strategies on synthetic data.

@@

\subsection{
    Interpolating synthetic survey with gradient boosting
}
\label{sec:gb_interpolation}
